Calculating Entropies With Nested Sampling

### 2017-02-27 paper/ms.tex
Provisional results

I consider two possible observing strategies, both involving
$n=101$ measurements. The `even' strategy has us observe at times
\begin{align}
t_i = \frac{i-1}{n-1}
\end{align}
for $i \in \{1, 2, ..., 101\}$, that is, the observation times are
$0, 0.01, 0.02, ..., 0.99, 1$ ---
evenly spaced from $t=0$ to $t=1$, including the endpoints.
The second, `uneven' observing strategy schedules the observations
according to
\begin{align}
t_i = \left(\frac{i - \frac{1}{2}}{n}\right)^3
\end{align}
which schedules observations close together initially, and the
spacing increases with time.

@@

Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for every quantity.
The results were
\begin{align}
H_{\tau}    &= 7.337 \pm 0.028 \\
H_{\d}      &= 903.8 \pm 1.1   \\
H_{\tau,\d} &= 900.8 \pm 1.1      % Interim result only
\end{align}

To convert these to differential entropies, we need to add the appropriate
log-volume to each. These are $-6.215$, $-790.32$, and
$-796.53$ respectively, so the differential entropies are:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 115.5 \pm 1.9   \\
H'_{\tau, \d} &= 105.7 \pm 1.7
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
uncertainties in each of these three quantities is quite large, and would lead
to a large uncertainty on $I_{\tau, \d}$ if computed naively. To reduce this
uncertainty, we computed the three entropies using a common random number
generator for the reference pairs. This is an example of the
{\em common random numbers} technique of variance reduction.
The resulting mutual information is
$I_{\tau, \d} = 10.35 \pm 0.61$ nats.

>>>>

I consider two possible observing strategies, both involving
$n=101$ measurements. The `even' strategy has observations at times
\begin{align}
t_i = \frac{i-1}{n-1}
\end{align}
for $i \in \{1, 2, ..., 101\}$, that is, the observation times are
$0, 0.01, 0.02, ..., 0.99, 1$ ---
evenly spaced from $t=0$ to $t=1$, including the endpoints.
The second, `uneven' observing strategy schedules the observations
according to
\begin{align}
t_i = \left(\frac{i - \frac{1}{2}}{n}\right)^3
\end{align}
which schedules observations close together initially, and further apart
as time goes on.

@@

Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for each of the three
runs. The {\todo interim} results are:
\begin{align}
H_{\tau}    &= 7.307 \pm 0.027 \\
H_{\d}      &= 905.5 \pm 5.6   \\
H_{\tau,\d} &= 883.6 \pm 5.2
\end{align}

To conver these to differential entropies, I applied the log-volume
correction factors for 1, 101, and 102 dimensions, which are
$-6.215$, $-790.32$, and $-798.62$ respectively. This gives
the following {\todo interim} values for the differential entropies:
\begin{align}
H'_{\tau}     &= 1.092 \pm 0.027 \label{eqn:h_tau}\\
H'_{\d}       &= 115.2 \pm 5.6   \\
H'_{\tau, \d} &=  93.3 \pm 5.2   \label{eqn:h_joint}
\end{align}
The mutual information is equal to $H_{\tau} + H_{\d} - H_{\tau, \d}$. The
uncertainties in each of these three quantities is quite large, and would lead
to a large uncertainty on $I_{\tau, \d}$ if computed from the results
in Equations~\ref{eqn:h_tau} to~\ref{eqn:h_joint}. To reduce this
uncertainty, we computed the three entropies using a common random number
generator for the reference pairs. This is an example of the
{\em common random numbers} technique of variance reduction.
The resulting {\todo provisional} mutual information is
$I_{\tau, \d} = 31.7 \pm 3.8$ nats.
